\bf The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

### 2021-08-07 paper/paper.Rmd
my edits are done!

We present a case-study of `forestecology`'s functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015]. The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US---the Blue Ridge, Ridge and Valley, and Piedmont provinces---and is adjacent to the northern end of Shenandoah National Park.

@@

We also load species information as it existed on GitHub on 2021/08/02, which includes family, genus, and species information; as well as classifications of the canopy position (canopy, canopy emergent, understory, shrub layer) and drought tolerance (intolerant, resistant) of the species.

@@

@@

@@

@@

Creative use of the grouping variable allows for tests of what best explains competitive interactions. Here we compare whether species idenity or canopy position of the species best explains.

@@

We find that species identity has a lower RMSE, so does a better job. But we can still plot the competition posteriors for the canopoy position groupings.

@@

>>>>

We present a case-study of `forestecology`'s functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015, @davies_forestgeo_2021]. The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US---the Blue Ridge, Ridge and Valley, and Piedmont provinces---and is adjacent to the northern end of Shenandoah National Park.

@@

We also load species information as it existed on GitHub on 2021/08/02, which includes family, genus, and species information; as well as classifications of the canopy position (canopy, canopy emergent, understory, shrub layer), drought tolerance (intolerant, resistant) and other characteristics of the species.

@@

Here we use basal area as the continuous predictor variable but the package is flexible to allow the user to specify any continuous predictor variable (basal area, biomass, tree height, a soil nutrient value, \dots). The package can also be used to compare predictor variables and see which best explains tree growth, see Appendix 3 \ref{compare-pred-var} for an example comparing basal area and above ground biomass. Similarly, the package can use any categorical variable as a predictor and compare between different categorical variables. For example in @allen_permutation_2020 we compare grouping individuals based on species, family, and based on trait-based groups. In Appendix 4 \ref{compare-group-var} we give another example and compare grouping individuals by species or by potential canopy position (canopy, understory, shrub layer).

@@

In the above code we use the basal area of an individual as a continuous predictor variable. But the package allows the user to specify any predictor variable in the `comp_x_var` argument of `create_focal_vs_comp` function. Here we use the cross-validated model comparison to see which of two possible predictor variables, basal area or above ground biomass, best explains growth.

@@

Here we see that basal area does a much better job than above ground biomass.

@@

The package also allows the user to specify the categorical predictor grouping variable. Here we compare two different such variables: species and the potential canopy position of that species. If we had individual-level crown classes (@smith_practice_1986 dominant, codominant, intermediate and suppressed) that could also be used.

@@

We find that species identity has a lower RMSE, so does a better job. But we can still plot the competition posteriors for the canopy position groupings.

@@

Unsurprisingly we see that canopy and canopy emergent competitors generally have negative effects on their neighbors, while shrubs and understory competitors have neutral or even positive effects.

### 2021-08-07 paper/paper.tex
my edits are done!

We present a case-study of \texttt{forestecology}'s functionality on
data from the Smithsonian Conservation Biology Institute (SCBI) large
forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO
global network of research sites
\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}.
The 25.6 ha (640 x 400 m) plot is located at the intersection of three
of the major physiographic provinces of the eastern US---the Blue Ridge,
Ridge and Valley, and Piedmont provinces---and is adjacent to the
northern end of Shenandoah National Park.

@@

@@

@@

Creative use of the grouping variable allows for tests of what best
explains competitive interactions. Here we compare whether species
idenity or canopy position of the species best explains.

>>>>

We present a case-study of \texttt{forestecology}'s functionality on
data from the Smithsonian Conservation Biology Institute (SCBI) large
forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO
global network of research sites
\citep[\citet{andersonteixeira_ctfs-forestgeo_2015},
\citet{davies_forestgeo_2021}]{bourg_initial_2013}. The 25.6 ha (640 x
400 m) plot is located at the intersection of three of the major
physiographic provinces of the eastern US---the Blue Ridge, Ridge and
Valley, and Piedmont provinces---and is adjacent to the northern end of
Shenandoah National Park.

@@

Here we use basal area as the continuous predictor variable but the
package is flexible to allow the user to specify any continuous
predictor variable (basal area, biomass, tree height, a soil nutrient
value, \dots). The package can also be used to compare predictor
variables and see which best explains tree growth, see Appendix 3
\ref{compare-pred-var} for an example comparing basal area and above
ground biomass. Similarly, the package can use any categorical variable
as a predictor and compare between different categorical variables. For
example in \citet{allen_permutation_2020} we compare grouping
individuals based on species, family, and based on trait-based groups.
In Appendix 4 \ref{compare-group-var} we give another example and
compare grouping individuals by species or by potential canopy position
(canopy, understory, shrub layer).

@@

In the above code we use the basal area of an individual as a continuous
predictor variable. But the package allows the user to specify any
predictor variable in the \texttt{comp\_x\_var} argument of
\texttt{create\_focal\_vs\_comp} function. Here we use the
cross-validated model comparison to see which of two possible predictor
variables, basal area or above ground biomass, best explains growth.

@@

The package also allows the user to specify the categorical predictor
grouping variable. Here we compare two different such variables: species
and the potential canopy position of that species. If we had
individual-level crown classes (\citet{smith_practice_1986} dominant,
codominant, intermediate and suppressed) that could also be used.
